---
title: "HW6"
author: "Sarahy Martinez"
date: "2024-12-01"
output: github_document
---

```{r}
library(tidyverse)
library(broom)
library(modelr)
set.seed(1)

```



#Problem 1 

```{r, eval=FALSE}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())


```



# Problem 2 

## Problem 2A

```{r}
#reading in Washington post data 

homicide_raw_data = read.csv("./homicide-data.csv")

homicide = homicide_raw_data %>% 
  unite(city_state, city, state, sep = ",") %>% 
  mutate(disposition_case = case_when( disposition == "Closed without arrest"~ "unresolved",
                                  disposition == "Closed by arrest" ~ "resolved",
                                  disposition == "Open/No arrest" ~ "unresolved")) %>%
 filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL")) %>% 
  filter(victim_race %in% c("White", "Black")) %>% 
  mutate(victim_age = as.numeric(victim_age)) %>% 
  mutate(resolved = as.numeric(disposition_case == "resolved"))
  


```

## Problem 2B 

Using these data, we can fit a logistic regression for the binary “resolved” outcome and victim demographics as predictors. This uses the glm function with the family specified to account for the non-Gaussian outcome distribution.

```{r}
baltimore_df = 
  homicide %>% 
  filter(city_state == "Baltimore,MD") %>% 
  mutate(
    resolved = as.numeric(disposition_case == "resolved"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>% 
  select(resolved,victim_age, victim_race, victim_sex)

# applying the glm function
fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

#using the broom tidy function and estimate the odds ratio and confidence interval

Balt_fit_logistic = 
  broom::tidy(fit_logistic, conf.int = TRUE) %>% #conf.int function used to calculate the CI
  mutate(
    OR = exp(estimate),
    OR_conf.low = exp(conf.low),
    OR_conf.high = exp(conf.high)
  ) %>%
  select (OR, OR_conf.low, OR_conf.high) %>% 
  rename(low = "OR_conf.low",
          high = "OR_conf.high") 

Balt_fit_logistic %>% knitr::kable(digits = 2)



```



## Problem 2C 

```{r}


# taking df homicide and nesting data by city
homicide_data = homicide %>%
  group_by(city_state) %>%
  nest()

# create a function for the logistic regression and extract OR and CI
logistic = function(data) {
  fit = glm(resolved ~ victim_age + victim_race + victim_sex, data = data, family = binomial())
  tidy_fit = tidy(fit, conf.int = TRUE) %>%
    mutate(
      OR = exp(estimate),
      OR_conf.low = exp(conf.low),
      OR_conf.high = exp(conf.high)
    ) %>%
    filter(term == "victim_sexMale") %>%
    select(term, log_OR = estimate, OR, p.value, OR_conf.low, OR_conf.high)
  return(tidy_fit)
}

# Apply the function to each city's data and unnest the results to return values
city_results =
  homicide_data %>%
  mutate(model = purrr::map(data, logistic)) %>%
  unnest(model) %>% 
    select (OR, OR_conf.low, OR_conf.high) %>% 
  rename(Low = "OR_conf.low",
          High = "OR_conf.high") 

# Display the results
city_results %>% knitr::kable(digits = 2)


```



## Problem 2D

```{r}
# plotting the estimated ORs and CIs for each city according to estimated OR 

city_estimated = ggplot(city_results, aes(x = reorder(city_state, OR), y =OR))+
                          geom_point()+
                          geom_line()+
                          geom_errorbar(aes(ymin = Low, ymax = High), width = 0.2)+
                          labs(title = "Estimated Odds Ratio and CI for Homcides Across Cities",
                                                       x = "City",
                                                        y = "Odds Ratio")+
                              theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Based on the plot we can see that New Mexico has the largest odds ratio  and confidence interval in comparison to New York who has the smallest OR and CI. This is interesting given that it is common knowledge that Chicago, IL has been considered the city and state with the highest homicide rate yet its OR is less than New York and Baton Rogue, LA. 

# Problem 3 

## Problem 3A 

```{r}

child_weight = read.csv("./birthweight.csv")

child_weight = child_weight %>% 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = fct_relevel(factor(frace, levels = c(1, 2, 3, 4, 8, 9),
                               labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
                        "White", "Black"),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    mrace = fct_relevel(factor(mrace, levels = c(1, 2, 3, 4, 8),
                               labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
                        "White", "Black"),
gaweeks = as.numeric(gaweeks))

```


## Problem 3b

```{r}

# developing the regression model, writing the linear regression model with all the potential predictors of low birth weight 

fit = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain, data = child_weight)

fitted.values(fit) # record of fitted values 
residuals(fit) # gives us the residuals 

# creating residuals and predictions 
modelr::add_residuals(child_weight, fit) 
modelr::add_predictions(child_weight, fit)

# adding the residuals and predictions to the df 
child_weight = child_weight %>% 
  add_predictions(fit) %>% 
  add_residuals(fit) 

# plotting 
 ggplot(child_weight, aes(x = pred, y = resid)) +
  geom_point()+
  geom_smooth()+
  labs(
    title = "Residuals vs. Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

```


## Question 3c

fitting the models and others 

```{r}
# original model 
fit_original = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain, data = child_weight)

# second model with only gestational age and length at birth as predictors 

fit_gest_length = lm(bwt ~ blength +gaweeks, data = child_weight)

# third model only using head circumference, length, sex, and interactions 

fit_interactions = lm(bwt~babysex +bhead + blength + babysex*bhead*blength,data = child_weight)

```

Comparing models using cross validations 

```{r}
# create cross validation df 

cv_df = 
  crossv_mc(child_weight, 100) %>%  #make validation set n =100
  mutate(
    train = map(train, as_tibble),
    test = map(train, as_tibble)
  )

# fit model to splits and calculate RMSE
cv_df = cv_df %>% 
  mutate(
    fit_original = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain, data = .x)), 
    fit_gest_length = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    fit_interactions = map(train, ~lm(bwt ~ babysex + bhead + blength + babysex * bhead * blength, data = .x))
  ) %>% 
  mutate(
    rmse_original = map2_dbl(fit_original, test, ~rmse(model = .x, data = .y)),
    rmse_gest_length = map2_dbl(fit_gest_length, test, ~rmse(model = .x, data = .y)),
    rmse_interactions = map2_dbl(fit_interactions, test, ~rmse(model = .x, data = .y))
  )


```


```{r}

# plotting the rmse's 

rmse_models = cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse))+
  geom_violin()+
  labs( title = "RMSE Distribution Across Cross-Validation Splits", 
        x = "Model", y = "RMSE" ) + 
  theme_minimal()


print(rmse_models)

```


The plot above includes three models the original, gestation age and length and the interactions. The graph shows that gest_length has the highest RMSE values meanwhile the original model with all predictors is has the least. This means that the gest_length performs the worst in terms of prediction accuracy. This also indicates that the interactions model is neither the worst or the best but the original model works the best in prediction accuracy since it has lower rmse values. The distribution for the original model is also more narrow than the interactions and gest_length which allows us to conclude that the original model is more consistent than the others. 


