---
title: "HW6"
author: "Sarahy Martinez"
date: "2024-12-01"
output: github_document
---

```{r}
library(tidyverse)
library(broom)
set.seed(1)

```



#Problem 1 

```{r, eval=FALSE}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())


```



# Problem 2 

## Problem 2A

```{r}
#reading in Washington post data 

homicide_raw_data = read.csv("./homicide-data.csv")

homicide = homicide_raw_data %>% 
  unite(city_state, city, state, sep = ",") %>% 
  mutate(disposition_case = case_when( disposition == "Closed without arrest"~ "unresolved",
                                  disposition == "Closed by arrest" ~ "resolved",
                                  disposition == "Open/No arrest" ~ "unresolved")) %>%
 filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL")) %>% 
  filter(victim_race %in% c("White", "Black")) %>% 
  mutate(victim_age = as.numeric(victim_age)) %>% 
  mutate(resolved = as.numeric(disposition_case == "resolved"))
  


```

## Problem 2B 

Using these data, we can fit a logistic regression for the binary “resolved” outcome and victim demographics as predictors. This uses the glm function with the family specified to account for the non-Gaussian outcome distribution.

```{r}
baltimore_df = 
  homicide %>% 
  filter(city_state == "Baltimore,MD") %>% 
  mutate(
    resolved = as.numeric(disposition_case == "resolved"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>% 
  select(resolved,victim_age, victim_race, victim_sex)

# applying the glm function
fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

#using the broom tidy function and estimate the odds ratio and confidence interval

Balt_fit_logistic = 
  broom::tidy(fit_logistic, conf.int = TRUE) %>% #conf.int function used to calculate the CI
  mutate(
    OR = exp(estimate),
    OR_conf.low = exp(conf.low),
    OR_conf.high = exp(conf.high)
  ) %>%
  select (OR, OR_conf.low, OR_conf.high) %>% 
  rename(low = "OR_conf.low",
          high = "OR_conf.high") 

Balt_fit_logistic %>% knitr::kable(digits = 2)



```



## Problem 2C 

```{r}


# taking df homicide and nesting data by city
homicide_data = homicide %>%
  group_by(city_state) %>%
  nest()

# create a function for the logistic regression and extract OR and CI
logistic = function(data) {
  fit = glm(resolved ~ victim_age + victim_race + victim_sex, data = data, family = binomial())
  tidy_fit = tidy(fit, conf.int = TRUE) %>%
    mutate(
      OR = exp(estimate),
      OR_conf.low = exp(conf.low),
      OR_conf.high = exp(conf.high)
    ) %>%
    filter(term == "victim_sexMale") %>%
    select(term, log_OR = estimate, OR, p.value, OR_conf.low, OR_conf.high)
  return(tidy_fit)
}

# Apply the function to each city's data and unnest the results to return values
city_results =
  homicide_data %>%
  mutate(model = purrr::map(data, logistic)) %>%
  unnest(model) %>% 
    select (OR, OR_conf.low, OR_conf.high) %>% 
  rename(Low = "OR_conf.low",
          High = "OR_conf.high") 

# Display the results
city_results %>% knitr::kable(digits = 2)


```



## Problem 2D

```{r}
# plotting the estimated ORs and CIs for each city according to estimated OR 

city_estimated = ggplot(city_results, aes(x = reorder(city_state, OR), y =OR))+
                          geom_point()+
                          geom_line()+
                          geom_errorbar(aes(ymin = Low, ymax = High), width = 0.2)+
                          labs(title = "Estimated Odds Ratio and CI for Homcides Across Cities",
                                                       x = "City",
                                                        y = "Odds Ratio")+
                              theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Based on the plot we can see that New Mexico has the largest odds ratio  and confidence interval in comparison to New York who has the smallest OR and CI. This is interesting given that it is common knowledge that Chicago, IL has been considered the city and state with the highest homicide rate yet its OR is less than New York and Baton Rogue, LA. 

# Problem 3 

## Problem 3A 

```{r}

child_weight = read.csv("./birthweight.csv")

child_weight = child_weight %>% 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = fct_relevel(factor(frace, levels = c(1, 2, 3, 4, 8, 9),
                               labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
                        "White", "Black"),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    mrace = fct_relevel(factor(mrace, levels = c(1, 2, 3, 4, 8),
                               labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
                        "White", "Black"),
gaweeks = as.numeric(gaweeks))

```


## Problem 3b

```{r}

# developing the regression model, writing the linear regression model with all the potential predictors of low birth weight 

fit = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain, data = child_weight)

fitted.values(fit) # record of fitted values 
residuals(fit) # gives us the residuals 

```











